---
title: "scGraphVerse Case Study: B-cell GRN Reconstruction"
author: "Francesco Cecere"
output:
  BiocStyle::html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{scGraphVerse Case Study: B-cell GRN Reconstruction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Introduction

This vignette demonstrates the **scGraphVerse** workflow on a single-cell 
RNA-seq dataset. We show how to:

1. Load and preprocess public PBMC data.
2. Infer gene regulatory networks with **GENIE3**.
3. Build consensus networks and detect communities.
4. Validate inferred edges using STRINGdb.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.width = 6,
    fig.height = 5
)
library(scGraphVerse)
library(Seurat)
library(tidyverse)
library(TENxPBMCData)
library(scater)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(SingleR)
library(celldex)
```

# 1. Dataset and Preprocessing

In this real-data analysis, we'll work with two public PBMC (Peripheral Blood 
Mononuclear Cell) datasets from 10X Genomics. Our strategy is to focus 
specifically on B cells and identify a common set of highly expressed genes 
across both datasets. This approach allows us to compare regulatory networks 
between different experimental batches while controlling for cell type and gene
selection effects.

**Data Processing Workflow:**
1. Load two PBMC datasets (3k and 4k cells) from TENxPBMCData
2. Use SingleR for automated cell type annotation
3. Select top 500 most expressed genes in B cells from each dataset
4. Find intersection of expressed genes to ensure comparable gene sets
5. Subset datasets to B cells only with common gene set

This preprocessing ensures we have clean, comparable data for network inference.

```{r loaddata}
# 1. Load PBMC data
pbmc_obj <- TENxPBMCData("pbmc3k")
pbmc_obj1 <- TENxPBMCData("pbmc4k")

sce <- logNormCounts(pbmc_obj)
symbols_tenx <- rowData(sce)$Symbol_TENx
valid <- !is.na(symbols_tenx) & symbols_tenx != ""
sce <- sce[valid, ]
rownames(sce) <- make.unique(symbols_tenx[valid])
logcounts(sce) <- as.matrix(logcounts(sce))
colnames(sce) <- paste0("cell_", seq_len(ncol(sce)))

sce1 <- logNormCounts(pbmc_obj1)
symbols_tenx1 <- rowData(sce1)$Symbol_TENx
valid1 <- !is.na(symbols_tenx1) & symbols_tenx1 != ""
sce1 <- sce1[valid1, ]
rownames(sce1) <- make.unique(symbols_tenx1[valid1])
logcounts(sce1) <- as.matrix(logcounts(sce1))
colnames(sce1) <- paste0("cell_", seq_len(ncol(sce1)))

# 2. Cell type annotation
ref <- celldex::HumanPrimaryCellAtlasData()
pred1 <- SingleR(test = sce1, ref = ref, labels = ref$label.main)
colData(sce1)$predicted_celltype <- pred1$labels

pred <- SingleR(test = sce, ref = ref, labels = ref$label.main)
colData(sce)$predicted_celltype <- pred$labels

# 3. Select top 500 B-cell genes
genes <- selgene(
    object = sce,
    top_n = 500,
    cell_type = "B_cell",
    cell_type_col = "predicted_celltype",
    remove_rib = TRUE,
    remove_mt = TRUE
)

genes1 <- selgene(
    object = sce1,
    top_n = 500,
    cell_type = "B_cell",
    cell_type_col = "predicted_celltype",
    remove_rib = TRUE,
    remove_mt = TRUE
)

# 4. Find intersection
commong <- intersect(genes, genes1)


# 5. Subset data
pbmc1_sub <- subset(sce, features = commong)
pbmc2_sub <- subset(sce1, features = commong)

pbmc1_sub <- pbmc1_sub[commong, ]
pbmc2_sub <- pbmc2_sub[commong, ]

pbmc1_sub <- pbmc1_sub[, colData(pbmc1_sub)$predicted_celltype == "B_cell"]
pbmc2_sub <- pbmc2_sub[, colData(pbmc2_sub)$predicted_celltype == "B_cell"]

# List for multi-sample analysis
bcell_list <- list(pbmc1_sub, pbmc2_sub)

```

# 2. Network Inference

Now we'll infer gene regulatory networks from our preprocessed B cell data 
using GENIE3. Unlike the simulation study where we had control over all 
parameters, here we're working with real biological data that presents unique 
challenges: B cells have distinct expression patterns, lower total gene counts
compared to the full transcriptome, and batch effects between the two PBMC
datasets.

```{r genie3}
# Choose method: "GENIE3", "GRNBoost2", "ZILGM", "PCzinb" or "JRF"
method <- "GENIE3"
networks <- infer_networks(
    count_matrices_list = bcell_list,
    method = method,
    nCores = 1
)

```

## 2.1. Building Adjacency Matrices

Here we apply a more stringent threshold (99th percentile) compared to the 
simulation study (95th percentile). This is appropriate for real data analysis
where we expect higher noise levels and want to focus on the most confident 
regulatory relationships. With B cell data, we're particularly interested in 
high-confidence edges controlling B cell function.

```{r adjacency, fig.alt="plot Graphs"}
# Weighted adjacency
wadj <- generate_adjacency(networks)
# Symmetrize
swadj <- symmetrize(wadj, weight_function = "mean")
# Binary cutoff (top 1%)
binary_adj <- cutoff_adjacency(
    count_matrices = bcell_list,
    weighted_adjm_list = swadj,
    n = 2,
    method = method,
    quantile_threshold = 0.99,
    nCores = 1
)
# Plot
plotg(binary_adj)
```

# 3. Consensus and Community Detection

With only two PBMC datasets, our consensus network will include edges that 
appear in both batches, providing high confidence in the regulatory 
relationships. This is different from the simulation study where we had three 
conditions and used a majority vote.

```{r consensus, fig.alt="plot consensus Graph"}
# Consensus by vote
consensus <- create_consensus(binary_adj, method = "vote")

plotg(list(consensus))
```


Here we demonstrate two key features not shown in the simulation study: 
**network comparison with external databases** and the `false_plot` 
parameter in `compare_consensus()`.

```{r comparecommunity, fig.alt="comparing Graphs with STRINGdb"}
# Note: This requires internet connection for STRINGdb
# To run this section, set eval=TRUE and ensure internet connectivity
compare_consensus(consensus_matrix = consensus, false_plot = TRUE)

# Community detection
communities <- community_path(consensus)
```

# 4. Validation with STRINGdb

Unlike the simulation study where we used STRINGdb to create ground truth, here 
we use it for validation of our inferred B cell regulatory network. 
The `keep_all_genes = TRUE` parameter ensures we retain information for all 
genes in our dataset, even those not found in STRING, which is important for 
comprehensive edge mining analysis.

```{r stringdb}
# Note: This requires internet connection for STRINGdb downloads
# To run this section, set eval=TRUE and ensure internet connectivity
str <- stringdb_adjacency(
    genes = rownames(consensus),
    species = 9606,
    required_score = 900,
    keep_all_genes = TRUE
)$binary
ground_truth <- symmetrize(list(str), weight_function = "mean")[[1]]

# Edge mining: TP rates
em <- edge_mining(list(consensus),
    query_edge_types = "TP",
    ground_truth = ground_truth
)
```

# 5. Conclusion

This case study illustrates how **scGraphVerse** enables end-to-end
GRN reconstruction and validation in single-cell data. 
Users can swap inference algorithms, tune thresholds,
and incorporate external prior networks.

# Session Information

```{r sessioninfo}
sessionInfo()
```

