---
title: "scGraphVerse Case Study: B-cell GRN Reconstruction"
author: "Francesco Cecere"
output:
  BiocStyle::html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{scGraphVerse Case Study: B-cell GRN Reconstruction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Introduction

This vignette demonstrates the **scGraphVerse** workflow on a single-cell 
RNA-seq dataset. We show how to:

1. Load and preprocess public PBMC data.
2. Infer gene regulatory networks with**GENIE3**.
3. Build consensus networks and detect communities.
4. Validate inferred edges using STRINGdb.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.width = 6,
    fig.height = 5
)
library(scGraphVerse)
library(Seurat)
library(tidyverse)
library(TENxPBMCData)
library(scater)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(SingleR)
library(celldex)
```

# 1. Dataset and Preprocessing

We use two public PBMC SCE objects containing PBMC samples.
Our goal is to focus on healthy B cells and a common gene set.

```{r loaddata}
# 1. Load PBMC data
pbmc_obj <- TENxPBMCData("pbmc3k")
pbmc_obj1 <- TENxPBMCData("pbmc4k")

sce <- logNormCounts(pbmc_obj)
symbols_tenx <- rowData(sce)$Symbol_TENx
valid <- !is.na(symbols_tenx) & symbols_tenx != ""
sce <- sce[valid, ]
rownames(sce) <- make.unique(symbols_tenx[valid])
logcounts(sce) <- as.matrix(logcounts(sce))
colnames(sce) <- paste0("cell_", seq_len(ncol(sce)))

sce1 <- logNormCounts(pbmc_obj1)
symbols_tenx1 <- rowData(sce1)$Symbol_TENx
valid1 <- !is.na(symbols_tenx1) & symbols_tenx1 != ""
sce1 <- sce1[valid1, ]
rownames(sce1) <- make.unique(symbols_tenx1[valid1])
logcounts(sce1) <- as.matrix(logcounts(sce1))
colnames(sce1) <- paste0("cell_", seq_len(ncol(sce1)))

ref <- celldex::HumanPrimaryCellAtlasData()
pred1 <- SingleR(test = sce1, ref = ref, labels = ref$label.main)
colData(sce1)$predicted_celltype <- pred1$labels

pred <- SingleR(test = sce, ref = ref, labels = ref$label.main)
colData(sce)$predicted_celltype <- pred$labels

# 2. Select top 500 B-cell genes
genes <- selgene(
    object = sce,
    top_n = 500,
    cell_type = "B_cell",
    cell_type_col = "predicted_celltype",
    remove_rib = TRUE,
    remove_mt = TRUE
)

# 2. Select top 500 B-cell genes
genes1 <- selgene(
    object = sce1,
    top_n = 500,
    cell_type = "B_cell",
    cell_type_col = "predicted_celltype",
    remove_rib = TRUE,
    remove_mt = TRUE
)

commong <- intersect(genes, genes1)

pbmc1_sub <- subset(sce, features = commong)
pbmc2_sub <- subset(sce1, features = commong)

pbmc1_sub <- pbmc1_sub[commong, ]
pbmc2_sub <- pbmc2_sub[commong, ]

pbmc1_sub <- pbmc1_sub[, colData(pbmc1_sub)$predicted_celltype == "B_cell"]
pbmc2_sub <- pbmc2_sub[, colData(pbmc2_sub)$predicted_celltype == "B_cell"]

# List for multi-sample analysis
bcell_list <- list(pbmc1_sub, pbmc2_sub)

shuffle_matrix_rows <- function(mat) {
    # Convert DelayedArray to matrix if needed
    if (is(mat, "DelayedArray")) {
        mat <- as.matrix(mat)
    }
    shuffled <- t(apply(mat, 1, sample))
    rownames(shuffled) <- rownames(mat)
    colnames(shuffled) <- colnames(mat)
    return(shuffled)
}



check <- shuffle_matrix_rows(assay(bcell_list[[1]], "counts")[1:5, 1:5])
```

# 2. Network Inference

We infer GRNs using three algorithms. Replace `method` to switch.

```{r genie3}
# Choose method: "GENIE3", "GRNBoost2", or "JRF"
method <- "GENIE3"
networks <- infer_networks(
    count_matrices_list = bcell_list,
    method = method,
    nCores = 1
)
```

## 2.1. Building Adjacency Matrices

Convert edge lists to weighted matrices, symmetrize, and apply threshold.

```{r adjacency, fig.alt="plot Graphs"}
# Weighted adjacency
wadj <- generate_adjacency(networks)
# Symmetrize
swadj <- symmetrize(wadj, weight_function = "mean")
# Binary cutoff (top 5%)
binary_adj <- cutoff_adjacency(
    count_matrices = bcell_list,
    weighted_adjm_list = swadj,
    n = 2,
    method = method,
    quantile_threshold = 0.95,
    nCores = 1
)
# Plot
plotg(binary_adj)
```

# 3. Consensus and Community Detection

Aggregate multiple binary networks into a consensus and find network modules.

```{r consensus, fig.alt="plot consensus Graph"}
# Consensus by vote
consensus <- create_consensus(binary_adj, method = "vote")
plotg(list(consensus))
```


```{r comparecommunity, fig.alt="comparing Graphs with STRINGdb", eval=FALSE}
# Note: This requires internet connection for STRINGdb
# To run this section, set eval=TRUE and ensure internet connectivity
compare_consensus(consensus_matrix = consensus, false_plot = TRUE)

# Community detection
communities <- community_path(consensus)
```

# 4. Validation with STRINGdb

Fetch high-confidence interactions from STRING and evaluate true positives.

```{r stringdb, eval=FALSE}
# Note: This requires internet connection for STRINGdb downloads
# To run this section, set eval=TRUE and ensure internet connectivity
str <- stringdb_adjacency(
    genes = rownames(consensus),
    species = 9606,
    required_score = 900,
    keep_all_genes = TRUE
)$binary
ground_truth <- symmetrize(list(str), weight_function = "mean")[[1]]

# Edge mining: TP rates
em <- edge_mining(list(consensus),
    query_edge_types = "TP",
    ground_truth = ground_truth
)
```

# 5. Conclusion

This case study illustrates how **scGraphVerse** enables end-to-end
GRN reconstruction and validation in single-cell data. 
Users can swap inference algorithms, tune thresholds,
and incorporate external prior networks.

# Session Information

```{r sessioninfo}
sessionInfo()
```

