---
title: "scGraphVerse Case Study: Zero-Inflated Simulation and GRN Inference"
author: "Francesco Cecere"
output:
  BiocStyle::html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{scGraphVerse Simulation Study: Sim & GRN Reconstruction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

scGraphVerse is a modular and extensible R package for constructing, comparing, 
and visualizing gene regulatory networks (GRNs) from single-cell RNAseq data.
It includes a collection of inference algorithms, post-processing utilities,
and visualization functions, all designed with interoperability, 
reproducibility, and scalability in mind.

Single-cell data presents unique challenges: high sparsity, batch variation, 
and the need to distinguish shared versus condition-specific regulation. 
scGraphVerse helps address these through multi-method inference, consensus
building, simulation benchmarking, and communities and pathway analysis. 
It starts with SingleCellExperiment, Seurat or matrix-based objects, enabling 
flexible pipelines across diverse experimental setups.

## Why use scGraphVerse?

While several GRN inference packages exist in Bioconductor 
(e.g. GENIE3, minet, SCENIC, netZoo), most are limited to single-method 
inference or lack support for multi-condition, multi-replicate, and 
multi-method comparisons. scGraphVerse adds value by:

1. Supporting multiple inference algorithms, including GENIE3, GRNBoost2, 
ZILGM, PCzinb, and Joint Random Forests (JRF).
2. Providing joint modeling of GRNs across multiple conditions via JRF
3. Enabling benchmarking simulations from known ground-truth GRNs with 
customizable ZINB-based count matrix generation.
4. Offering community analysis, including consensus GRNs, similarity analysis,
and STRINGdb integration.

Users can plug in their single-cell objects, infer networks and visualize
regulatory structure using built-in radar plots, ROC curves, 
and community overlays.

## How scGraphVerse works

The typical scGraphVerse workflow begins with one or more count matrices, 
or objects like SingleCellExperiment or Seurat. These are passed to the core 
function infer_networks(), which runs the selected inference methods and
returns an adjacency matrix or list of matrices.

For benchmarking or teaching purposes, synthetic datasets can be generated 
using zinb_simdata() based on a known adjacency matrix. This supports method
comparison across early, late, and joint integration strategies using 
earlyj() and compare_consensus().

An end-to-end workflow, from network inference to community detection and 
consensus visualization, is provided in the rest of vignettes.

As intended for both high-throughput analysis and intuitive usage, 
scGraphVerse supports custom workflows using its modular functions.

# Simulation study 

In this simulation study, we use **scGraphVerse** to:

1. Define a ground-truth regulatory network from high-confidence interactions.
2. Simulate zero-inflated scRNA-seq count data that respects the ground truth.
3. Infer gene regulatory networks using **GENIE3**.
4. Evaluate performance with ROC curves, precision–recall scores, 
and community similarity.
5. Build consensus networks and perform edge mining.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.width = 6,
    fig.height = 5
)
library(scGraphVerse)
library(igraph)
library(ggraph)
library(TENxPBMCData)
library(scater)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(SingleR)
library(celldex)
```

# 1. Defining a Ground-Truth Network from STRINGdb

We select 500 top-variable T-cell genes and fetch high-confidence edges 
(score ≥ 900) from STRINGdb as our ground truth.

```{r, eval=FALSE}
# Note: This section requires internet connectivity for external data downloads
# To run this section, set eval=TRUE and ensure internet connectivity

# 1. Load PBMC data
pbmc_obj <- TENxPBMCData("pbmc3k")

sce <- logNormCounts(pbmc_obj)
symbols_tenx <- rowData(sce)$Symbol_TENx
valid <- !is.na(symbols_tenx) & symbols_tenx != ""
sce <- sce[valid, ]
rownames(sce) <- make.unique(symbols_tenx[valid])
logcounts(sce) <- as.matrix(logcounts(sce))
colnames(sce) <- paste0("cell_", seq_len(ncol(sce)))

ref <- celldex::HumanPrimaryCellAtlasData()
pred <- SingleR(test = sce, ref = ref, labels = ref$label.main)
colData(sce)$predicted_celltype <- pred$labels

# 2. Select top 500 T-cell genes
genes <- selgene(
    object = sce,
    top_n = 500,
    cell_type = "T_cells",
    cell_type_col = "predicted_celltype",
    remove_rib = TRUE,
    remove_mt = TRUE
)

# 3. Retrieve STRINGdb adjacency
str_res <- stringdb_adjacency(
    genes = genes,
    species = 9606,
    required_score = 900,
    keep_all_genes = FALSE
)

wadj_truth <- str_res$weighted
adj_truth <- str_res$binary
adj_truth <- adj_truth[order(rownames(adj_truth)), order(colnames(adj_truth))]

# 5. Visualize network
gtruth <- graph_from_adjacency_matrix(adj_truth, mode = "undirected")
ggraph(gtruth, layout = "fr") +
    geom_edge_link(color = "gray") +
    geom_node_point(color = "steelblue") +
    ggtitle(paste0(
        "Ground Truth: ",
        vcount(gtruth),
        " nodes, ",
        ecount(gtruth),
        " edges"
    )) +
    theme_minimal()
```

For demonstration purposes, we'll use the built-in example data:

```{r demo-data, fig.alt="Ground truth network nodes and edges"}
# Use built-in example data for demonstration
data("adj_truth")
data("count_matrices")

# Visualize the example ground truth network
gtruth <- igraph::graph_from_adjacency_matrix(adj_truth, mode = "undirected")
ggraph(gtruth, layout = "fr") +
    geom_edge_link(color = "gray") +
    geom_node_point(color = "steelblue") +
    ggtitle(paste0(
        "Example Ground Truth: ",
        igraph::vcount(gtruth),
        " nodes, ",
        igraph::ecount(gtruth),
        " edges"
    )) +
    theme_minimal()
```

# 2. Simulating Zero-Inflated Count Data

We simulate three batches (n=50 cells each) of count matrices that follow the 
ground-truth network topology with dropout.

```{r simulate}
# Simulation parameters
nodes <- nrow(adj_truth)
sims <- zinb_simdata(
    n = 50,
    p = nodes,
    B = adj_truth,
    mu_range = list(c(1, 4), c(1, 7), c(1, 10)),
    mu_noise = c(1, 3, 5),
    theta = c(1, 0.7, 0.5),
    pi = c(0.2, 0.2, 0.2),
    kmat = 3,
    depth_range = c(0.8 * nodes * 3, 1.2 * nodes * 3)
)
# Transpose to cells × genes
count_matrices <- lapply(sims, t)
```

# 3. Inferring Networks with JRF

We run **GENIE3** across the simulated batches to infer batch-specific
regulatory edges.

```{r genie3}
networks_joint <- infer_networks(
    count_matrices_list = count_matrices,
    method = "JRF",
    nCores = 15
)
# Weighted adjacency
wadj_list <- generate_adjacency(networks_joint)
# Symmetrize weights
swadj_list <- symmetrize(wadj_list, weight_function = "mean")
```

# 4. ROC Curve and AUC

Plot the ROC curve comparing continuous edge weights to the binary ground truth.

```{r roc, fig.alt="ROC curve"}
roc_res <- plotROC(
    swadj_list,
    adj_truth,
    plot_title = "ROC Curve: JRF Joint Integration",
    is_binary = FALSE
)
roc_res$plot
auc_joint <- roc_res$auc
```

## 4.1. Precision–Recall and Graph Visualization

Compute precision scores and visualize the binary networks.

```{r cutoff, fig.alt="result Graphs"}
# Binary cutoff at 95th percentile
binary_listj <- cutoff_adjacency(
    count_matrices = count_matrices,
    weighted_adjm_list = swadj_list,
    n = 2,
    method = "JRF",
    quantile_threshold = 0.95,
    nCores = 15,
    debug = TRUE
)

# Precision scores
pscores_joint <- pscores(adj_truth, binary_listj)
head(pscores_joint)

# Network plot
plotg(binary_listj)
```

# 5. Consensus Networks and Community Similarity

Aggregate inferred binaries by majority vote and compare community structure
to ground truth.

```{r consensus, fig.alt="consensus Graph"}
# Consensus matrix
consensus <- create_consensus(binary_listj, method = "vote")

plotg(list(consensus))
```


```{r communities adj_truth, fig.alt="Community detection adj_truth"}
# Note: This section may require internet connectivity for STRINGdb validation
# To run this section, set eval=TRUE and ensure internet connectivity

# Compare consensus to truth
evaluate_consensus <- compare_consensus(
    consensus_matrix = consensus,
    reference_matrix = adj_truth,
    false_plot = FALSE
)

# Community detection

comm_truth <- community_path(adj_truth)
```


```{r communities, fig.alt="Community detection consensus"}
comm_cons <- community_path(consensus)
```


```{r communities metrics, fig.alt="compare Graph with reference"}
# Similarity
sim_score <- community_similarity(comm_truth, list(comm_cons))
```

## 5.1. Edge Mining

Identify true positive edges in the consensus network using edge mining.

```{r edgemining}
em <- edge_mining(list(consensus), adj_truth, query_edge_types = "TP")
```

```{r sessioninfo}
sessionInfo()
```

