---
title: "scGraphVerse Case Study: Zero-Inflated Simulation and GRN Inference"
author: "Francesco Cecere"
output:
  BiocStyle::html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{scGraphVerse Simulation Study: Sim & GRN Reconstruction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Simulation study 

In this case study, we use **scGraphVerse** to:

1. Define a ground-truth regulatory network from high-confidence interactions.
2. Simulate zero-inflated scRNA-seq count data that respects the ground truth.
3. Infer gene regulatory networks using **GENIE3**.
4. Evaluate performance with ROC curves, precision–recall scores, 
and community similarity.
5. Build consensus networks and perform edge mining.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.width = 6,
    fig.height = 5
)
library(scGraphVerse)
library(igraph)
library(ggraph)
```

# 1. Defining a Ground-Truth Network from STRINGdb

We select 500 top-variable T-cell genes and fetch high-confidence edges 
(score ≥ 900) from STRINGdb as our ground truth.

```{r ground-truth, fig.alt="Reference graph"}
# 1. Download PBMC data
url <- paste0(
    "https://www.dropbox.com/s/r8qwsng79rhp9gf/",
    "SCA_scRNASEQ_TISSUE_WHOLE_BLOOD.RDS?dl=1"
)
seu <- download_Atlas(file_url = url)

# 2. Select top 500 T-cell genes
genes <- selgene(
    object = seu,
    top_n = 500,
    cell_type = "T_cells",
    cell_type_col = "CELL_TYPE",
    remove_rib = TRUE,
    remove_mt = TRUE
)

# 3. Retrieve STRINGdb adjacency
str_res <- stringdb_adjacency(
    genes = genes,
    species = 9606,
    required_score = 900,
    keep_all_genes = FALSE
)
wadj_truth <- str_res$weighted
adj_truth <- str_res$binary
adj_truth <- adj_truth[order(rownames(adj_truth)), order(colnames(adj_truth))]

# 5. Visualize network
gtruth <- graph_from_adjacency_matrix(adj_truth, mode = "undirected")
ggraph(gtruth, layout = "fr") +
    geom_edge_link(color = "gray") +
    geom_node_point(color = "steelblue") +
    ggtitle(paste0(
        "Ground Truth: ",
        vcount(gtruth),
        " nodes, ",
        ecount(gtruth),
        " edges"
    )) +
    theme_minimal()
```

# 2. Simulating Zero-Inflated Count Data

We simulate three batches (n=50 cells each) of count matrices that follow the 
ground-truth network topology with dropout.

```{r simulate-zilgm}
# Simulation parameters
nodes <- nrow(adj_truth)
sims <- zinb_simdata(
    n = 50,
    p = nodes,
    B = adj_truth,
    mu_range = list(c(1, 4), c(1, 7), c(1, 10)),
    mu_noise = c(1, 3, 5),
    theta = c(1, 0.7, 0.5),
    pi = c(0.2, 0.2, 0.2),
    kmat = 3,
    depth_range = c(0.8 * nodes * 3, 1.2 * nodes * 3)
)
# Transpose to cells × genes
count_matrices <- lapply(sims, t)

```

# 3. Inferring Networks with JRF

We run **Joint Random Forests** across the simulated batches to infer shared 
and batch-specific regulatory edges.

```{r infer-jrf}
networks_joint <- infer_networks(
    count_matrices_list = count_matrices,
    method = "GENIE3",
    nCores = 10
)
# Weighted adjacency
wadj_list <- generate_adjacency(networks_joint)
# Symmetrize weights
swadj_list <- symmetrize(wadj_list, weight_function = "mean")
```

# 4. ROC Curve and AUC

Plot the ROC curve comparing continuous edge weights to the binary ground truth.

```{r roc-curve, fig.alt="ROC curve"}
roc_res <- plotROC(
    swadj_list,
    adj_truth,
    plot_title = "ROC Curve: JRF Joint Integration",
    is_binary = FALSE
)
roc_res$plot
auc_joint <- roc_res$auc
```

## 4.1. Precision–Recall and Graph Visualization

Compute precision scores and visualize the binary networks.

```{r pr-and-plot, fig.alt="result Graphs"}
# Binary cutoff at 95th percentile
binary_listj <- cutoff_adjacency(
    count_matrices = count_matrices,
    weighted_adjm_list = swadj_list,
    n = 2,
    method = "GENIE3",
    quantile_threshold = 0.95,
    nCores = 10,
    debug = TRUE
)

# Precision scores
pscores_joint <- pscores(adj_truth, binary_listj)
head(pscores_joint)

# Network plot
plotg(binary_listj)
```

# 5. Consensus Networks and Community Similarity

Aggregate inferred binaries by majority vote and compare community structure
to ground truth.

```{r consensus, fig.alt="consensus Graph"}
# Consensus matrix
consensus <- create_consensus(binary_listj, method = "vote")
plotg(list(consensus))
```


```{r communities, "compare Graph with reference}
# Compare consensus to truth
evaluate_consensus <- compare_consensus(
    consensus_matrix = consensus,
    reference_matrix = adj_truth,
    false_plot = FALSE
)

# Community detection
comm_truth <- community_path(adj_truth)
comm_cons <- community_path(consensus)
# Similarity
sim_score <- community_similarity(comm_truth, list(comm_cons))
```

## 5.1. Edge Mining

Identify true positive edges in the consensus network using edge mining.

```{r edge-mining}
em <- edge_mining(list(consensus), adj_truth, query_edge_types = "TP")
```

```{r session-info}
sessionInfo()
```

