---
title: "PBMC analysis"
author: "Francesco Cecere"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: cerulean
    df_print: paged
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r}
knitr::opts_knit$set(root.dir = "/home/francescoc/Desktop/scGraphVerse/data/",message=FALSE, warning=FALSE)
```

## Load Libraries and data

```{r}
library(RColorBrewer)
library(DT)
library(tidyverse)
library(Seurat)
library(STRINGdb)
library(SingleCellExperiment)
library(scGraphVerse)
library(BiocParallel)
library(GENIE3)

reticulate::use_python("/usr/bin/python3", required = TRUE)
arboreto <- reticulate::import("arboreto.algo")
pandas <- reticulate::import("pandas")
numpy <- reticulate::import("numpy")

time <- list()

seurat_subset_filt <- readRDS("./../data/PBMC.top1200.RDS")

Seurat::GetAssayData(seurat_split, assay = "RNA", slot = "data")
```

```{r}
seurat_split <- SplitObject(
  object = seurat_subset_filt,
  split.by = "donor_id"
)
```

## GENIE3

### Late integration

```{r}
#sc_split <- lapply(seurat_split, as.SingleCellExperiment)

set.seed(1234)
time[["GENIE3_late_15Cores"]] <- system.time(
  genie3_late <- infer_networks(seurat_split, 
                                method="GENIE3",
                                nCores = 15)
)

#saveRDS(genie3_late, paste(ddir, "genie3_late_pbmc1000.RDS", sep = "/"))
```

#### Symmetrize and ROC

```{r}
genie3_late_wadj <- generate_adjacency(genie3_late)
sgenie3_late_wadj <- symmetrize(genie3_late_wadj, weight_function = "mean")

#genie3_late_auc <- plotROC(sgenie3_late_wadj, adj, plot_title = "ROC curve - GENIE3 Late Integration")

```

#### Cutoff

```{r}
sgenie3_late_adj <- cutoff_adjacency(count_matrices = seurat_split,
                                     weighted_adjm_list = sgenie3_late_wadj, 
                                     n = 3,
                                     method = "GENIE3",
                                     nCores = 15)

#scores.genie3.late.all <- pscores(adj, sgenie3_late_adj)
plots <- plotg(sgenie3_late_adj)

#cbind(scores.genie3.late.all$Statistics, genie3_late_auc) %>% write.table(., paste(ddir, "genie3_late_pbmc1000_perf.txt", sep = "/"), sep = "\t", quote = F, col.names = T, row.names = F)
```

#### Consensus

```{r}
consesusm <- create_consensus(sgenie3_late_adj, method="vote")
consesusu <- create_consensus(sgenie3_late_adj, method="union")
consesunet <- create_consensus(adj_matrix_list = sgenie3_late_adj, weighted_list = sgenie3_late_wadj, method = "INet", threshold = 0.05, ncores = 15)

plotg(list(consesusm))
plotg(list(consesusu))
plotg(list(consesunet))

#scores.genie3.late <- pscores(adj, list(consesusm))
#scoresu.genie3.late <- pscores(adj, list(consesusu))
#scoresnet.genie3.late <- pscores(adj, list(consesunet))

```

#### Plot comparison

```{r}
adj <- stringdb_adjacency(
  genes          = rownames(sgenie3_late_adj[[1]]),
  species        = 9606,
  required_score = 900,
  keep_all_genes = T
)

adj <- adj$binary
adj <- adj[rownames(consesusm),rownames(consesusm)]
adj <- symmetrize(list(adj), weight_function = "mean")
adj <- adj[[1]]

plotg(list(adj))

ajm_compared <- compare_consensus(consesusm, adj)
ajm_compared <- compare_consensus(consesusu, adj)
ajm_compared <- compare_consensus(consesunet, adj)

```

#### PubMed Query

```{r}
query <- edge_mining(list(consesusm), adj, query_edge_types = "TP")

query[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "TP PubMed query")

query <- edge_mining(list(consesusm), adj, query_edge_types = "FP")

query[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "FP PubMed query")

query <- edge_mining(list(consesusm), adj, query_edge_types = "FN")

query[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "FN PubMed query")
```

### Early integration

```{r}
early_matrix <- earlyj(seurat_split)

set.seed(1234)
time[["GENIE3_early_15Cores"]] <- system.time(
  genie3_early <- infer_networks(list(early_matrix), method="GENIE3", nCores = 15)
)

#saveRDS(genie3_early, paste(ddir, "genie3_early_pbmc1000.RDS", sep = "/"))
```

#### Symmetrize and ROC

```{r}
genie3_early_wadj <- generate_adjacency(genie3_early)
sgenie3_early_wadj <- symmetrize(genie3_early_wadj, weight_function = "mean")
#genie3_early_auc <- plotROC(sgenie3_early_wadj, adj, plot_title = "ROC curve - GENIE3 Early Integration")

```

#### Cutoff

```{r}
sgenie3_early_adj <- cutoff_adjacency(count_matrices = list(early_matrix),
                                      weighted_adjm_list = sgenie3_early_wadj, 
                                      n = 2,
                                      method = "GENIE3",
                                      nCores = 15)

#scores.genie3.early <- pscores(adj, sgenie3_early_adj)
plots <- plotg(sgenie3_early_adj)


#cbind(scores.genie3.early$Statistics, genie3_early_auc) %>% write.table(., paste(ddir, "genie3_early_pbmc1000_perf.txt", sep = "/"), sep = "\t", quote = F, col.names = T, row.names = F)
```

#### Plot comparison

```{r}
adj <- stringdb_adjacency(
  genes          = rownames(sgenie3_early_adj[[1]]),
  species        = 9606,
  required_score = 900,
  keep_all_genes = T
)

adj <- adj$binary
adj <- adj[rownames(sgenie3_early_adj[[1]]),rownames(sgenie3_early_adj[[1]])]
adj <- symmetrize(list(adj), weight_function = "mean")
adj <- adj[[1]]

plotg(list(adj))

ajm_compared <- compare_consensus(sgenie3_early_adj[[1]], adj)

```

#### PubMed Query

```{r}
query <- edge_mining(sgenie3_early_adj, adj, query_edge_types = "TP")

query[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "TP PubMed query")

query <- edge_mining(sgenie3_early_adj, adj, query_edge_types = "FP")

query[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "FP PubMed query")

query <- edge_mining(sgenie3_early_adj, adj, query_edge_types = "FN")

query[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "FN PubMed query")
```

## GRNBoost2

### Late integration

```{r}
set.seed(1234)
time[["grnboost_late"]] <- system.time(
  grnboost_late <- infer_networks(seurat_split, 
                                method="GRNBoost2",
                                nCores = 15)
)

#saveRDS(grnboost_late, paste(ddir, "grnboost_late_pbmc1000.RDS", sep = "/"))
```

#### Symmetrize and ROC

```{r}
grnboost_late_wadj <- generate_adjacency(grnboost_late)
sgrnboost_late_wadj <- symmetrize(grnboost_late_wadj, weight_function = "mean")
#grnboost_late_auc <- plotROC(sgrnboost_late_wadj, adj, plot_title = "ROC curve - grnboost Late Integration")

```

#### Cutoff

```{r}
sgrnboost_late_adj <- cutoff_adjacency(count_matrices = seurat_split,
                                     weighted_adjm_list = sgrnboost_late_wadj, 
                                     n = 3,
                                     method = "GRNBoost2",
                                     nCores = 15)

#scores.grnboost.late.all <- pscores(adj, sgrnboost_late_adj)
plots <- plotg(sgrnboost_late_adj)

#cbind(scores.grnboost.late.all$Statistics, grnboost_late_auc) %>% write.table(., paste(ddir, "grnboost_late_pbmc1000_perf.txt", sep = "/"), sep = "\t", quote = F, col.names = T, row.names = F)
```

#### Consensus

```{r}
consesusm <- create_consensus(sgrnboost_late_adj, method="vote")
consesusu <- create_consensus(sgrnboost_late_adj, method="union")
consesunet <- create_consensus(adj_matrix_list = sgrnboost_late_adj, weighted_list = sgrnboost_late_wadj, method = "INet", threshold = 0.05)

#scores.grnboost.late <- pscores(adj, list(consesusm))
#scoresu.grnboost.late <- pscores(adj, list(consesusu))
#scoresnet.grnboost.late <- pscores(adj, list(consesunet))


```

#### Plot comparison

```{r}
adj <- stringdb_adjacency(
  genes          = rownames(sgrnboost_late_adj[[1]]),
  species        = 9606,
  required_score = 900,
  keep_all_genes = T
)

adj <- adj$binary
adj <- adj[rownames(consesusm),rownames(consesusm)]
adj <- symmetrize(list(adj), weight_function = "mean")
adj <- adj[[1]]

plotg(list(adj))

ajm_compared <- compare_consensus(consesusm, adj)
ajm_compared <- compare_consensus(consesusu, adj)
ajm_compared <- compare_consensus(consesunet, adj)

```

#### PubMed Query

```{r}
query <- edge_mining(list(consesusm), adj, query_edge_types = "TP")

query[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "TP PubMed query")

query <- edge_mining(list(consesusm), adj, query_edge_types = "FP")

query[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "FP PubMed query")

query <- edge_mining(list(consesusm), adj, query_edge_types = "FN")

query[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "FN PubMed query")
```

### Early integration

```{r}
early_matrix <- earlyj(seurat_split)

set.seed(1234)
time[["grnboost_early"]] <- system.time(
  grnboost_early <- infer_networks(list(early_matrix), method="GRNBoost2")
)

#saveRDS(grnboost_early, paste(ddir, "grnboost_early_pbmc1000.RDS", sep = "/"))
```

#### Symmetrize and ROC

```{r}
grnboost_early_wadj <- generate_adjacency(grnboost_early)
sgrnboost_early_wadj <- symmetrize(grnboost_early_wadj, weight_function = "mean")
#grnboost_early_auc <- plotROC(sgrnboost_early_wadj, adj, plot_title = "ROC curve - grnboost Early Integration")

```

#### Cutoff

```{r}
sgrnboost_early_adj <- cutoff_adjacency(count_matrices = list(early_matrix),
                                      weighted_adjm_list = sgrnboost_early_wadj, 
                                      n = 2,
                                      method = "GRNBoost2",
                                      nCores = 15)

#scores.grnboost.early <- pscores(adj, sgrnboost_early_adj)
plots <- plotg(sgrnboost_early_adj)

#cbind(scores.grnboost.early$Statistics, grnboost_early_auc) %>% write.table(., paste(ddir, "grnboost_early_pbmc1000_perf.txt", sep = "/"), sep = "\t", quote = F, col.names = T, row.names = F)
```

#### Plot comparison

```{r}
adj <- stringdb_adjacency(
  genes          = rownames(sgrnboost_early_adj[[1]]),
  species        = 9606,
  required_score = 900,
  keep_all_genes = T
)

adj <- adj$binary
adj <- adj[rownames(sgrnboost_early_adj[[1]]),rownames(sgrnboost_early_adj[[1]])]
adj <- symmetrize(list(adj), weight_function = "mean")
adj <- adj[[1]]

plotg(list(adj))

ajm_compared <- compare_consensus(sgrnboost_early_adj[[1]], adj)
```

#### PubMed Query

```{r}
query <- edge_mining(sgrnboost_early_adj, adj, query_edge_types = "TP")

query[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "TP PubMed query")

query <- edge_mining(sgrnboost_early_adj, adj, query_edge_types = "FP")

query[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "FP PubMed query")

query <- edge_mining(sgrnboost_early_adj, adj, query_edge_types = "FN")

query[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "FN PubMed query")

```


## Joint Integration

### Joint Random Forest

```{r JRF inference}
#https://cran.r-project.org/src/contrib/Archive/JRF/
#install.packages("/home/francescoc/Downloads/JRF_0.1-4.tar.gz", repos = NULL, type = "source")
set.seed(1234)
time[["JRF"]] <- system.time(
  jrf_mat <- infer_networks(seurat_split, method="JRF", nCores = 15)
)
 
jrf_mat[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "JRF output")

```

#### Prepare the output

```{r}
jrf_list <- list()

importance_columns <- grep("importance", names(jrf_mat[[1]]), value = TRUE)

for (i in seq_along(importance_columns)) {
# Select the 'gene1', 'gene2', and the current 'importance' column
  df <- jrf_mat[[1]][, c("gene1", "gene2", importance_columns[i])]
  
  # Rename the importance column to its original name (e.g., importance1, importance2, etc.)
  names(df)[3] <- importance_columns[i]
  
  # Add the data frame to the output list
  jrf_list[[i]] <- df
}

saveRDS(jrf_list, "./../analysis/jrf.RDS")

jrf_list[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "JRF output")
```

#### symmetrize Output and ROC

```{r JRF symmetrizeROC late}
jrf_wadj <- generate_adjacency(jrf_list)
sjrf_wadj <- symmetrize(jrf_wadj, weight_function = "mean")

sjrf_wadj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "JRF symmetrize output")
```

#### Generate Adjacency and Apply Cutoff

```{r JRF cutoff late}
sjrf_adj <- cutoff_adjacency(count_matrices = count_matrices,
                 weighted_adjm_list = sjrf_wadj, 
                 n = 3,
                 method = "JRF")

sjrf_adj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "JRF adjacency")

```


```{r JRF consensus and score late}
consesusm <- create_consensus(sjrf_adj, method="vote")
consesusu <- create_consensus(sjrf_adj, method="union")
consesunet <- create_consensus(adj_matrix_list = sjrf_adj, weighted_list = sjrf_wadj, method = "INet", threshold = 0.1, ncores = 15)
```

```{r JRF compare late, fig.width=8,fig.width=10}
ajm_compared <- compare_consensus(consesusm, adj)
ajm_compared <- compare_consensus(consesusu, adj)
ajm_compared <- compare_consensus(consesunet, adj)
```

#### PubMed Query

```{r}
query <- edge_mining(list(consesusm), adj, query_edge_types = "TP")

query[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "TP PubMed query")

query <- edge_mining(list(consesusm), adj, query_edge_types = "FP")

query[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "FP PubMed query")

query <- edge_mining(list(consesusm), adj, query_edge_types = "FN")

query[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "FN PubMed query")
```

