---
title: "scGRN testing"
author: "Francesco Cecere"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: cerulean
    df_print: paged
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

# Flowchart

```{r setup, include=FALSE}
knitr::opts_chunk$set("/home/francescoc/Desktop/scGRN_simulation/R")
```

```{r global options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.height=5, fig.width=10, fig.align = "center", class.source = "foldable", timing = TRUE)
```

```{r libraries, timing = TRUE}
library(GENIE3)
library(doParallel)
library(igraph)
library(tidyverse)
library(DT)
library(reticulate)
library(learn2count)
library(rbenchmark)
library(reshape2)
library(gridExtra)
library(DiagrammeR)
library(pROC)
library(JRF)
library(DiagrammeRsvg)
library(rsvg)
library(RColorBrewer)
library(rbenchmark)

use_python("/usr/bin/python3", required = TRUE)
arboreto <- import("arboreto.algo")
pandas <- import("pandas")
numpy <- import("numpy")

```

```{r myfunc}
source("generate_adjacency.R")
source("symmetrize.R")
source("pscores.R")
source("plotg.R")
source("compare_consensus.R")
source("create_consensus.R")
source("earlyj.R")
source("plotROC.R")
source("cutoff_adjacency.R")
source("infer_networks.R")
```

```{r flowchart, fig.width=9, fig.height=15}
grViz_output <- DiagrammeR::grViz("
digraph biological_workflow {
  # Set up the graph attributes
  graph [layout = dot, rankdir = TB]

  # Define consistent node styles
  node [shape = rectangle, style = filled, color = lightblue, fontsize = 12]

  # Define nodes for each step
  StartNode [label = 'Ground Thruth - String Regulatory Network', shape = oval, color = seagreen, fontcolor = black]
  AdjacencyMatrix [label = 'Thruth Adjacency Matrix', shape = rectangle, color = seagreen]
  SimulateData [label = 'Simulate Single-Cell Data', shape = rectangle, color = goldenrod]

  # Reconstruction using Three Packages
  LateIntegration [label = 'Late\nIntegration', shape = oval, color = khaki]
  EarlyIntegration [label = 'Early\nIntegration', shape = oval, color = khaki]
  Jointanalysis [label = 'Joint\nanalysis', shape = oval, color = khaki]
  

  # Process 
  earlyj [label = 'earlyj.R', shape=diamond, color=lightblue, fontcolor=black]
  networkinference [label = 'infer_networks.R\nGENIE3\nGRNBoost2\nJRF', shape = rectangle, color = goldenrod, fontcolor=black]
  symmetrize [label = 'symmetrize.R', shape = rectangle, color = goldenrod, fontcolor=black]
  plotROC [label = 'plotROC.R', shape=diamond, color=lightblue, fontcolor=black]
  generateadjacency [label='generate_adjacency.R\nWeighted Adjacency', shape=rectangle, color=goldenrod, fontcolor=black]
  cutoffadjacency [label='cutoff_adjacency.R\nBinary Adjacency', shape=rectangle, color=goldenrod, fontcolor=black]
  pscores [label='pscores.R\nTPR\nFPR\nF1\nAccuracy\nPrecision', shape=diamond, color=lightblue, fontcolor=black]
  voting [label='Edges voting', shape=diamond, color=lightblue, fontcolor=black]
  plotgcompare  [label='plotg.R\ncompare_consesus.R\nPlot Graphs', shape=rectangle, color=goldenrod, fontcolor=black]

  # Define the workflow structure
  StartNode -> AdjacencyMatrix
  AdjacencyMatrix -> SimulateData
  SimulateData -> LateIntegration
  SimulateData -> EarlyIntegration
  SimulateData -> Jointanalysis
  EarlyIntegration -> earlyj
  earlyj -> networkinference
  LateIntegration -> networkinference
  Jointanalysis -> networkinference
  networkinference -> symmetrize
  symmetrize -> plotROC
  symmetrize -> generateadjacency
  generateadjacency -> cutoffadjacency
  cutoffadjacency -> pscores
  cutoffadjacency -> voting
  voting -> plotgcompare
  voting -> pscores
}
")

svg_code <- export_svg(grViz_output)
rsvg::rsvg_png(charToRaw(svg_code), "./../analysis/flowchart.png")

grViz_output
```

# Tcell Ground Truth

```{r readGT, fig.width=5, fig.height=5}
adjm <- read.table("./../data/adjacency_matrix.csv", header = T, row.names = 1, sep = ",") %>% as.matrix()
diag(adjm) <- 0

adjm %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "Ground Truth")
```


```{r plotGT, fig.width=5, fig.height=5}
gtruth <- igraph::graph_from_adjacency_matrix(adjm, mode = "undirected", diag = F)

num_nodes <- vcount(gtruth)
num_edges <- ecount(gtruth)

set.seed(1234)
plot(gtruth, 
     main = paste("Ground Truth\nNodes:", num_nodes, "Edges:", num_edges),
     vertex.label.color = "black",
     vertex.size = 6, 
     edge.width = 2, 
     vertex.label = NA,
     vertex.color = "steelblue",
     layout = igraph::layout_with_fr)
```

# Simulate Data

```{r simulate matrices}
ncell <- 500
nodes <- nrow(adjm)

set.seed(1130)
mu_values <- c(3, 5, 7)

count_matrices <- lapply(1:3, function(i) {
  set.seed(1130 + i)
  mu_i <- mu_values[i]
  
  count_matrix_i <- simdata(n = ncell, p = nodes, B = adjm, family = "ZINB", 
                            mu = mu_i, mu_noise = 1, theta = 0.5, pi = 0.2)
  
  count_matrix_df <- as.data.frame(count_matrix_i)
  colnames(count_matrix_df) <- colnames(adjm)
  rownames(count_matrix_df) <- paste("cell", 1:nrow(count_matrix_df), sep = "")
  
  return(count_matrix_df)
})

count_matrices[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "Simulated count matrix")

saveRDS(count_matrices, "./../analysis/count_matrices.RDS")
```

# Matrices Integration
## Late Integration

### GENIE3

```{r GENIE3 late}
set.seed(1234)
genie3_late <- infer_networks(count_matrices, method="GENIE3")
saveRDS(genie3_late, "./../analysis/genie3_late.RDS")

genie3_late[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GENIE3 output")
```

#### symmetrize Output and ROC

```{r GENIE3 symmetrizeROC late}
genie3_late_wadj <- generate_adjacency(genie3_late, ground.truth = adjm)
sgenie3_late_wadj <- symmetrize(genie3_late_wadj, weight_function = "mean")
plotROC(sgenie3_late_wadj, adjm, plot_title = "ROC curve - GENIE3 Late Integration")

sgenie3_late_wadj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GENIE3 symmetrize output")
```

#### Generate Adjacency and Apply Cutoff

```{r GENIE3 cutoff late}
source("cutoff_adjacency.R")
sgenie3_late_adj <- cutoff_adjacency(count_matrices = count_matrices,
                 weighted_adjm_list = sgenie3_late_wadj, 
                 ground.truth = adjm,
                 n = 2,
                 method = "GENIE3")

sgenie3_late_adj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GENIE3 adjacency")

```

#### Comparison with the Ground Truth

```{r GENIE3 metrics late}
scores <- pscores(adjm, sgenie3_late_adj)

scores$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
```


```{r GENIE3 plotg late}
plots <- plotg(sgenie3_late_adj)
```

```{r GENIE3 consensus and score late}
consesusm <- create_consensus(sgenie3_late_adj, method="vote")
consesusu <- create_consensus(sgenie3_late_adj, method="union")

par(mfrow = c(2, 1))
scores <- pscores(adjm, list(consesusm))
scoresu <- pscores(adjm, list(consesusu))
par(mfrow = c(1, 1))

scores$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
scoresu$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")

```


```{r GENIE3 compare late, fig.width=8,fig.width=10}
ajm_compared <- compare_consensus(consesusm, adjm)
ajm_compared <- compare_consensus(consesusu, adjm)
```

### GRNBoost2

```{r GRNBoost2 late}
set.seed(1234)
grnb_late <- infer_networks(count_matrices, method="GRNBoost2")
saveRDS(grnb_late, "./../analysis/grnb_late.RDS")

grnb_late[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GRNBoost2 output")
```

#### symmetrize Output and ROC

```{r GRNBoost2 symmetrizeROC late}
grnb_late_wadj <- generate_adjacency(grnb_late, ground.truth = adjm)
sgrnb_late_wadj <- symmetrize(grnb_late_wadj, weight_function = "mean")
plotROC(sgrnb_late_wadj, adjm, plot_title = "ROC curve - GRNBoost2 Late Integration")

sgrnb_late_wadj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GRNBoost2 symmetrize output")
```

#### Generate Adjacency and Apply Cutoff

```{r GRNBoost2 cutoff late}
sgrnb_late_adj <- cutoff_adjacency(count_matrices = count_matrices,
                 weighted_adjm_list = sgrnb_late_wadj, 
                 ground.truth = adjm,
                 n = 2,
                 method = "GRNBoost2")

sgrnb_late_adj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GRNBoost2 adjacency")

```

#### Comparison with the Ground Truth

```{r GRNBoost2 metrics late}
scores <- pscores(adjm, sgrnb_late_adj)

scores$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
```


```{r GRNBoost2 plotg late}
plots <- plotg(sgrnb_late_adj)
```

```{r GRNBoost2 consensus and score late}
consesusm <- create_consensus(sgrnb_late_adj, method="vote")
consesusu <- create_consensus(sgrnb_late_adj, method="union")

par(mfrow = c(2, 1))
scores <- pscores(adjm, list(consesusm))
scoresu <- pscores(adjm, list(consesusu))
par(mfrow = c(1, 1))

scores$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
scoresu$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
```

```{r GRNBoost2 compare late, fig.width=8,fig.width=10}
ajm_compared <- compare_consensus(consesusm, adjm)
ajm_compared <- compare_consensus(consesusu, adjm)
```


## Early Integration

```{r earlym}
early_matrix <- list(earlyj(count_matrices))
```

### GENIE3

```{r run GENIE3 early}
set.seed(1234)
genie3_early <- infer_networks(early_matrix, method="GENIE3")

saveRDS(genie3_early, "./../analysis/genie3_early.RDS")

genie3_early[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GENIE3 output")
```


#### symmetrize Output and ROC

```{r GENIE3 symmetrizeROC early}
genie3_early_wadj <- generate_adjacency(genie3_early, ground.truth = adjm)
sgenie3_early_wadj <- symmetrize(genie3_early_wadj, weight_function = "mean")
plotROC(sgenie3_early_wadj, adjm, plot_title = "ROC curve - GENIE3 Early Integration")

sgenie3_early_wadj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GENIE3 symmetrize output")
```

#### Generate Adjacency and Apply Cutoff

```{r GENIE3 cutoff early}
sgenie3_early_adj <- cutoff_adjacency(count_matrices = early_matrix,
                 weighted_adjm_list = sgenie3_early_wadj, 
                 ground.truth = adjm,
                 n = 2,
                 method = "GENIE3")

sgenie3_early_adj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GENIE3 adjacency")

```

#### Comparison with the Ground Truth

```{r GENIE3 metrics early}
scores <- pscores(adjm, sgenie3_early_adj)

scores$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
```


```{r GENIE3 plotg early}
plots <- plotg(sgenie3_early_adj)
```

```{r GENIE3 compare early, fig.width=8,fig.width=10}
ajm_compared <- compare_consensus(sgenie3_early_adj[[1]], adjm)
```


### GRNBoost2

```{r GRNBoost2 early}
set.seed(1234)
grnb_early <- infer_networks(early_matrix, method="GRNBoost2")
saveRDS(grnb_early, "./../analysis/grnb_early.RDS")

grnb_early[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GRNBoost2 output")
```

#### symmetrize Output and ROC

```{r GRNBoost2 symmetrizeROC early}
grnb_early_wadj <- generate_adjacency(grnb_early, ground.truth = adjm)
sgrnb_early_wadj <- symmetrize(grnb_early_wadj, weight_function = "mean")
plotROC(sgrnb_early_wadj, adjm, plot_title = "ROC curve - GRNBoost2 Early Integration")

grnb_early_wadj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GRNBoost2 symmetrize output")
```

#### Generate Adjacency and Apply Cutoff

```{r GRNBoost2 cutoff early}
sgrnb_early_adj <- cutoff_adjacency(count_matrices = early_matrix,
                 weighted_adjm_list = sgrnb_early_wadj, 
                 ground.truth = adjm,
                 n = 2,
                 method = "GRNBoost2")

sgrnb_early_adj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GRNBoost2 adjacency")

```

#### Comparison with the Ground Truth

```{r GRNBoost2 metrics early}
scores <- pscores(adjm, sgrnb_early_adj)

scores$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
```


```{r GRNBoost2 plotg early}
plots <- plotg(sgrnb_early_adj)
```

```{r GRNBoost2 compare early, fig.width=8,fig.width=10}
ajm_compared <- compare_consensus(sgrnb_early_adj[[1]], adjm)
```


## Joint Integration

### Joint Random Forest


```{r}
#https://cran.r-project.org/src/contrib/Archive/JRF/
#install.packages("/home/francescoc/Downloads/JRF_0.1-4.tar.gz", repos = NULL, type = "source")
#jrf_mat <- infer_networks(count_matrices, method="JRF")

#jrf_matrices <- lapply(count_matrices, t)
#jrf_matrices_norm <- lapply(jrf_matrices,function(x) {
#  (x - mean(x)) / sd(x)
#  })

#genes <- rownames(jrf_matrices_norm[[1]])
      
#netout <- JRF(X = jrf_matrices_norm, 
#              genes.name = genes, 
#              ntree = 500, 
#              mtry = round(sqrt(length(genes) - 1)))

#netout %>%
#    datatable(extensions = 'Buttons',
#              options = list(
#                dom = 'Bfrtip',
#                buttons = c('csv', 'excel'),
#                scrollX = TRUE,
#                pageLength = 10), 
#              caption = "JRF output")

#out.perm <- Run_permutation(jrf_matrices_norm,mtry=round(sqrt(length(genes)-1)),ntree=500, genes,3)
#out <- JRF_permutation(jrf_matrices_norm,mtry=round(sqrt(length(genes)-1)),ntree=500,genes,2)

#final.net <- JRF_network(netout,out.perm,0.001)
#final.net
```



```{r JRF inference}
#https://cran.r-project.org/src/contrib/Archive/JRF/
#install.packages("/home/francescoc/Downloads/JRF_0.1-4.tar.gz", repos = NULL, type = "source")
jrf_mat <- infer_networks(count_matrices, method="JRF")

jrf_mat[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "JRF output")

```

#### Prepare the output

```{r}
jrf_list <- list()

importance_columns <- grep("importance", names(jrf_mat[[1]]), value = TRUE)

for (i in seq_along(importance_columns)) {
  # Select the 'gene1', 'gene2', and the current 'importance' column
  df <- jrf_mat[[1]][, c("gene1", "gene2", importance_columns[i])]
  
  # Rename the importance column to its original name (e.g., importance1, importance2, etc.)
  names(df)[3] <- importance_columns[i]
  
  # Add the data frame to the output list
  jrf_list[[i]] <- df
}

saveRDS(jrf_list, "./../analysis/jrf.RDS")

jrf_list[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "JRF output")
```

#### symmetrize Output and ROC

```{r JRF symmetrizeROC late}
jrf_wadj <- generate_adjacency(jrf_list, ground.truth = adjm)
sjrf_wadj <- symmetrize(jrf_wadj, weight_function = "mean")
plotROC(sjrf_wadj, adjm, plot_title = "ROC curve - JRF Late Integration")

sjrf_wadj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "JRF symmetrize output")
```

#### Generate Adjacency and Apply Cutoff

```{r JRF cutoff late}
sjrf_adj <- cutoff_adjacency(count_matrices = count_matrices,
                 weighted_adjm_list = sjrf_wadj, 
                 ground.truth = adjm,
                 n = 2,
                 method = "JRF")

sjrf_adj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "JRF adjacency")

```

#### Comparison with the Ground Truth

```{r JRF metrics late}
scores <- pscores(adjm, sjrf_adj)

scores$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
```


```{r JRF plotg late}
plots <- plotg(sjrfadj)
```

```{r JRF consensus and score late}
consesusm <- create_consensus(sjrf_adj, method="vote")
consesusu <- create_consensus(sjrf_adj, method="union")

scores <- pscores(adjm, list(consesusm))
scoresu <- pscores(adjm, list(consesusu))

scores$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
scoresu$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
```

```{r JRF compare late, fig.width=8,fig.width=10}
ajm_compared <- compare_consensus(consesusm, adjm)
ajm_compared <- compare_consensus(consesusu, adjm)
```

```{r}
sessionInfo()
```

