---
title: "scGRN testing"
author: "Francesco Cecere"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: cerulean
    df_print: paged
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

# Flowchart

```{r setup, include=FALSE}
knitr::opts_chunk$set("/home/francescoc/Desktop/scGRN_simulation/R")
```

```{r global options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.height=5, fig.width=10, fig.align = "center", class.source = "foldable", timing = TRUE)
```

```{r libraries, timing = TRUE}
library(GENIE3)
library(doParallel)
library(igraph)
library(tidyverse)
library(DT)
library(reticulate)
library(learn2count)
library(rbenchmark)
library(reshape2)
library(gridExtra)
library(DiagrammeR)
library(pROC)
library(JRF)
library(DiagrammeRsvg)
library(rsvg)
library(RColorBrewer)
library(rbenchmark)
library(ZILGM)

use_python("/usr/bin/python3", required = TRUE)
arboreto <- import("arboreto.algo")
pandas <- import("pandas")
numpy <- import("numpy")

execution_times <- list()
```

```{r myfunc}
source("generate_adjacency.R")
source("symmetrize.R")
source("pscores.R")
source("plotg.R")
source("compare_consensus.R")
source("create_consensus.R")
source("earlyj.R")
source("plotROC.R")
source("cutoff_adjacency.R")
source("infer_networks.R")
```

```{r flowchart, fig.width=9, fig.height=15}
grViz_output <- DiagrammeR::grViz("
digraph biological_workflow {
  # Set up the graph attributes
  graph [layout = dot, rankdir = TB]

  # Define consistent node styles
  node [shape = rectangle, style = filled, color = lightblue, fontsize = 12]

  # Define nodes for each step
  StartNode [label = 'Ground Thruth - String Regulatory Network', shape = oval, color = seagreen, fontcolor = black]
  AdjacencyMatrix [label = 'Thruth Adjacency Matrix', shape = rectangle, color = seagreen]
  SimulateData [label = 'Simulate Single-Cell Data', shape = rectangle, color = goldenrod]

  # Reconstruction using Three Packages
  LateIntegration [label = 'Late\nIntegration', shape = oval, color = khaki]
  EarlyIntegration [label = 'Early\nIntegration', shape = oval, color = khaki]
  Jointanalysis [label = 'Joint\nanalysis', shape = oval, color = khaki]
  

  # Process 
  earlyj [label = 'earlyj.R', shape=diamond, color=lightblue, fontcolor=black]
  networkinference [label = 'infer_networks.R\nGENIE3\nGRNBoost2\nZILGM\nJRF', shape = rectangle, color = goldenrod, fontcolor=black]
  symmetrize [label = 'symmetrize.R', shape = rectangle, color = goldenrod, fontcolor=black]
  plotROC [label = 'plotROC.R', shape=diamond, color=lightblue, fontcolor=black]
  generateadjacency [label='generate_adjacency.R\nWeighted Adjacency', shape=rectangle, color=goldenrod, fontcolor=black]
  cutoffadjacency [label='cutoff_adjacency.R\nBinary Adjacency', shape=rectangle, color=goldenrod, fontcolor=black]
  pscores [label='pscores.R\nTPR\nFPR\nF1\nAccuracy\nPrecision', shape=diamond, color=lightblue, fontcolor=black]
  voting [label='Voting\nUnion\nIntersection', shape=diamond, color=lightblue, fontcolor=black]
  plotgcompare  [label='plotg.R\ncompare_consesus.R\nPlot Graphs', shape=rectangle, color=goldenrod, fontcolor=black]

  # Define the workflow structure
  StartNode -> AdjacencyMatrix
  AdjacencyMatrix -> SimulateData
  SimulateData -> LateIntegration
  SimulateData -> EarlyIntegration
  SimulateData -> Jointanalysis
  EarlyIntegration -> earlyj
  earlyj -> networkinference
  LateIntegration -> networkinference
  Jointanalysis -> networkinference
  networkinference -> symmetrize
  symmetrize -> plotROC
  symmetrize -> generateadjacency
  generateadjacency -> cutoffadjacency
  cutoffadjacency -> pscores
  cutoffadjacency -> voting
  voting -> plotgcompare
  voting -> pscores
}
")

svg_code <- export_svg(grViz_output)
rsvg::rsvg_png(charToRaw(svg_code), "./../analysis/flowchart.png")

grViz_output
```

# Tcell Ground Truth

```{r readGT, fig.width=5, fig.height=5}
adjm <- read.table("./../data/adjacency_matrix.csv", header = T, row.names = 1, sep = ",") %>% as.matrix()
diag(adjm) <- 0

adjm %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "Ground Truth")
```


```{r plotGT, fig.width=5, fig.height=5}
gtruth <- igraph::graph_from_adjacency_matrix(adjm, mode = "undirected", diag = F)

num_nodes <- vcount(gtruth)
num_edges <- ecount(gtruth)

set.seed(1234)
plot(gtruth, 
     main = paste("Ground Truth\nNodes:", num_nodes, "Edges:", num_edges),
     vertex.label.color = "black",
     vertex.size = 6, 
     edge.width = 2, 
     vertex.label = NA,
     vertex.color = "steelblue",
     layout = igraph::layout_with_fr)
```

# Simulate Data

```{r simulate matrices}
ncell <- 500
nodes <- nrow(adjm)

set.seed(1130)
mu_values <- c(3, 6, 9)

count_matrices <- lapply(1:3, function(i) {
  set.seed(1130 + i)
  mu_i <- mu_values[i]
  
  count_matrix_i <- simdata(n = ncell, p = nodes, B = adjm, family = "ZINB", 
                            mu = mu_i, mu_noise = 1, theta = 0.5, pi = 0.2)
  
  count_matrix_df <- as.data.frame(count_matrix_i)
  colnames(count_matrix_df) <- colnames(adjm)
  rownames(count_matrix_df) <- paste("cell", 1:nrow(count_matrix_df), sep = "")
  
  return(count_matrix_df)
})

count_matrices[[1]] %>%
    datatable(extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('csv', 'excel'),
              scrollX = TRUE,
              pageLength = 10), 
            caption = "Simulated count matrix")

saveRDS(count_matrices, "./../analysis/count_matrices.RDS")
```

# Matrices Integration
## Late Integration

### GENIE3

```{r GENIE3 late}
set.seed(1234)
tictoc::tic("GENIE3 late")
genie3_late <- infer_networks(count_matrices, method="GENIE3")
saveRDS(genie3_late, "./../analysis/genie3_late.RDS")
execution_times[['GENIE3 late']] <- tictoc::toc(log = TRUE)$toc[[1]]

genie3_late[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GENIE3 output")
```

#### symmetrize Output and ROC

```{r GENIE3 symmetrizeROC late}
genie3_late_wadj <- generate_adjacency(genie3_late, ground.truth = adjm)
sgenie3_late_wadj <- symmetrize(genie3_late_wadj, weight_function = "mean")
plotROC(sgenie3_late_wadj, adjm, plot_title = "ROC curve - GENIE3 Late Integration")

sgenie3_late_wadj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GENIE3 symmetrize output")
```

#### Generate Adjacency and Apply Cutoff

```{r GENIE3 cutoff late}
sgenie3_late_adj <- cutoff_adjacency(count_matrices = count_matrices,
                 weighted_adjm_list = sgenie3_late_wadj, 
                 ground.truth = adjm,
                 n = 2,
                 method = "GENIE3")

sgenie3_late_adj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GENIE3 adjacency")

```

#### Comparison with the Ground Truth

```{r GENIE3 metrics late}
scores.genie3.late.all <- pscores(adjm, sgenie3_late_adj)

scores.genie3.late.all$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
```


```{r GENIE3 plotg late}
plots <- plotg(sgenie3_late_adj)
```

```{r GENIE3 consensus and score late}
consesusm <- create_consensus(sgenie3_late_adj, method="vote")
consesusu <- create_consensus(sgenie3_late_adj, method="union")

scores.genie3.late <- pscores(adjm, list(consesusm))
scoresu.genie3.late <- pscores(adjm, list(consesusu))

scores.genie3.late$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
scores.genie3.late$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")

```


```{r GENIE3 compare late, fig.width=8,fig.width=10}
ajm_compared <- compare_consensus(consesusm, adjm)
ajm_compared <- compare_consensus(consesusu, adjm)
```

### GRNBoost2

```{r GRNBoost2 late}
set.seed(1234)
tictoc::tic("GRNBoost2 late")
grnb_late <- infer_networks(count_matrices, method="GRNBoost2")
saveRDS(grnb_late, "./../analysis/grnb_late.RDS")
execution_times[['GRNBoost2 late']] <- tictoc::toc(log = TRUE)$toc[[1]]

grnb_late[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GRNBoost2 output")
```

#### symmetrize Output and ROC

```{r GRNBoost2 symmetrizeROC late}
grnb_late_wadj <- generate_adjacency(grnb_late, ground.truth = adjm)
sgrnb_late_wadj <- symmetrize(grnb_late_wadj, weight_function = "mean")
plotROC(sgrnb_late_wadj, adjm, plot_title = "ROC curve - GRNBoost2 Late Integration")

sgrnb_late_wadj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GRNBoost2 symmetrize output")
```

#### Generate Adjacency and Apply Cutoff

```{r GRNBoost2 cutoff late}
sgrnb_late_adj <- cutoff_adjacency(count_matrices = count_matrices,
                 weighted_adjm_list = sgrnb_late_wadj, 
                 ground.truth = adjm,
                 n = 2,
                 method = "GRNBoost2")

sgrnb_late_adj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GRNBoost2 adjacency")

```

#### Comparison with the Ground Truth

```{r GRNBoost2 metrics late}
scores.grn.late.all <- pscores(adjm, sgrnb_late_adj)

scores.grn.late.all$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
```


```{r GRNBoost2 plotg late}
plots <- plotg(sgrnb_late_adj)
```

```{r GRNBoost2 consensus and score late}
consesusm <- create_consensus(sgrnb_late_adj, method="vote")
consesusu <- create_consensus(sgrnb_late_adj, method="union")

scores.grn.late <- pscores(adjm, list(consesusm))
scoresu.grn.late <- pscores(adjm, list(consesusu))

scores.grn.late$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
scoresu.grn.late$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
```

```{r GRNBoost2 compare late, fig.width=8,fig.width=10}
ajm_compared <- compare_consensus(consesusm, adjm)
ajm_compared <- compare_consensus(consesusu, adjm)
```



### ZILGM Park

```{r ZILGM late}
source("infer_networks.R")
set.seed(1234)
tictoc::tic("ZILGM late")
zilgm_late <- infer_networks(count_matrices_list = count_matrices, method = "ZILGM")
saveRDS(zilgm_late, "./../analysis/zilgm_late.RDS")
execution_times[['ZILGM late']] <- tictoc::toc(log = TRUE)$toc[[1]]
```


#### Comparison with the Ground Truth



```{r ZILGM scores late}
scores.zilgm.late.all <- pscores(adjm, est_graphs)
scoresu.zilgm.late.all <- pscores(adjm, est_graphs)

scores.zilgm.late.all$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
scoresu.zilgm.late.all$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")

```


```{r}
plots <- plotg(est_graphs)
```


```{r ZILGM scores consensus}
consesusm <- create_consensus(est_graphs, method="vote")
consesusu <- create_consensus(est_graphs, method="union")

scores.zilgm.late <- pscores(adjm, list(consesusm))
scoresu.zilgm.late <- pscores(adjm, list(consesusu))

scores.zilgm.late$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
scoresu.zilgm.late$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
```

```{r ZILGM compare late, fig.width=8,fig.width=10}
ajm_compared <- compare_consensus(consesusm, adjm)
ajm_compared <- compare_consensus(consesusu, adjm)
```

## Early Integration

```{r earlym}
early_matrix <- list(earlyj(count_matrices))
```

### GENIE3

```{r run GENIE3 early}
set.seed(1234)
tictoc::tic("GENIE3 early")
genie3_early <- infer_networks(early_matrix, method="GENIE3")
execution_times[['GENIE3 early']] <- tictoc::toc(log = TRUE)$toc[[1]]

saveRDS(genie3_early, "./../analysis/genie3_early.RDS")

genie3_early[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GENIE3 output")
```

#### symmetrize Output and ROC

```{r GENIE3 symmetrizeROC early}
genie3_early_wadj <- generate_adjacency(genie3_early, ground.truth = adjm)
sgenie3_early_wadj <- symmetrize(genie3_early_wadj, weight_function = "mean")
plotROC(sgenie3_early_wadj, adjm, plot_title = "ROC curve - GENIE3 Early Integration")

sgenie3_early_wadj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GENIE3 symmetrize output")
```

#### Generate Adjacency and Apply Cutoff

```{r GENIE3 cutoff early}
sgenie3_early_adj <- cutoff_adjacency(count_matrices = early_matrix,
                 weighted_adjm_list = sgenie3_early_wadj, 
                 ground.truth = adjm,
                 n = 2,
                 method = "GENIE3")

sgenie3_early_adj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GENIE3 adjacency")

```

#### Comparison with the Ground Truth

```{r GENIE3 metrics early}
scores.genie3.early <- pscores(adjm, sgenie3_early_adj)

scores.genie3.early$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
```


```{r GENIE3 plotg early}
plots <- plotg(sgenie3_early_adj)
```

```{r GENIE3 compare early, fig.width=8,fig.width=10}
ajm_compared <- compare_consensus(sgenie3_early_adj[[1]], adjm)
```

### GRNBoost2

```{r GRNBoost2 early}
set.seed(1234)
tictoc::tic("GRNBoost2 early")
grnb_early <- infer_networks(early_matrix, method="GRNBoost2")
execution_times[['GRNBoost2 early']] <- tictoc::toc(log = TRUE)$toc[[1]]

saveRDS(grnb_early, "./../analysis/grnb_early.RDS")

grnb_early[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GRNBoost2 output")
```

#### symmetrize Output and ROC

```{r GRNBoost2 symmetrizeROC early}
grnb_early_wadj <- generate_adjacency(grnb_early, ground.truth = adjm)
sgrnb_early_wadj <- symmetrize(grnb_early_wadj, weight_function = "mean")
plotROC(sgrnb_early_wadj, adjm, plot_title = "ROC curve - GRNBoost2 Early Integration")

grnb_early_wadj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GRNBoost2 symmetrize output")
```

#### Generate Adjacency and Apply Cutoff

```{r GRNBoost2 cutoff early}
sgrnb_early_adj <- cutoff_adjacency(count_matrices = early_matrix,
                 weighted_adjm_list = sgrnb_early_wadj, 
                 ground.truth = adjm,
                 n = 2,
                 method = "GRNBoost2")

sgrnb_early_adj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "GRNBoost2 adjacency")

```

#### Comparison with the Ground Truth

```{r GRNBoost2 metrics early}
scores.grn.early <- pscores(adjm, sgrnb_early_adj)

scores.grn.early$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
```


```{r GRNBoost2 plotg early}
plots <- plotg(sgrnb_early_adj)
```

```{r GRNBoost2 compare early, fig.width=8,fig.width=10}
ajm_compared <- compare_consensus(sgrnb_early_adj[[1]], adjm)
```

### ZILGM Park

```{r ZILGM early}
set.seed(123)

tictoc::tic("ZILGM early")
lambda_max = find_lammax(as.matrix(early_matrix[[1]]))
lambda_min = 1e-4 * lambda_max
lambs = exp(seq(log(lambda_max), log(lambda_min), length.out = 50))

nb2_fit = zilgm(X = as.matrix(early_matrix[[1]]), lambda = lambs, family = "NBII", update_type = "IRLS", do_boot = TRUE,
                  boot_num = 10, sym = "OR", nCores = 15)

est_graph = nb2_fit$network[[nb2_fit$opt_index]]
execution_times[['ZILGM early']] <- tictoc::toc(log = TRUE)$toc[[1]]
```

```{r ZILGM early}
set.seed(123)

lambda_max = find_lammax(as.matrix(count_matrices[[1]]))
lambda_min = 1e-4 * lambda_max
lambs = exp(seq(log(lambda_max), log(lambda_min), length.out = 50))

nb2_fit = zilgm(X = as.matrix(count_matrices[[1]]), lambda = lambs, family = "NBII", update_type = "IRLS", do_boot = TRUE,
                  boot_num = 10, sym = "OR", nCores = 15)

est_graph = nb2_fit$network[[nb2_fit$opt_index]]
```


#### Comparison with the Ground Truth

```{r ZIGLM score early}
scores.zilgm.early <- pscores(adjm, list(as.matrix(est_graph)))

scores.zilgm.early$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")

```

```{r ZILGM compare early, fig.width=8,fig.width=10}
ajm_compared <- compare_consensus(as.matrix(est_graph), adjm)
```


## Joint Integration

### Joint Random Forest


```{r}
#https://cran.r-project.org/src/contrib/Archive/JRF/
#install.packages("/home/francescoc/Downloads/JRF_0.1-4.tar.gz", repos = NULL, type = "source")
#jrf_mat <- infer_networks(count_matrices, method="JRF")

jrf_matrices <- lapply(count_matrices, t)
jrf_matrices_norm <- lapply(jrf_matrices,function(x) {
  (x - mean(x)) / sd(x)
  })

genes <- rownames(jrf_matrices_norm[[1]])

set.seed(1234)
tictoc::tic("JRF_out")
netout <- JRF(X = jrf_matrices_norm, 
              genes.name = genes, 
              ntree = 500, 
              mtry = round(sqrt(length(genes) - 1)))
execution_times[['JRF_out']] <- tictoc::toc(log = TRUE)$toc[[1]]

netout %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "JRF output")

out.perm <- Run_permutation(jrf_matrices_norm,mtry=round(sqrt(length(genes)-1)),ntree=500, genes,3)
#out <- JRF_permutation(jrf_matrices_norm,mtry=round(sqrt(length(genes)-1)),ntree=500,genes,2)

final.net <- JRF_network(netout,out.perm,0.001)
final.net
```



#### Finalnet 

```{r}
myJRF_network <- function(out.jrf, out.perm, TH) {
  
  nclasses <- dim(out.perm)[3]
  M <- dim(out.perm)[2]
  out <- vector("list", nclasses)  # Initialize the list to store results for each class

  for (net in 1:nclasses) { 
    j.np <- sort(out.jrf[, 2 + net], decreasing = TRUE)
    FDR <- matrix(0, dim(out.perm)[1], 1)
    
    th <- NULL  # Initialize threshold variable
    for (s in 1:length(j.np)) { 
      FP <- sum(sum(out.perm[, , net] >= j.np[s])) / M
      FDR[s] <- FP / s
      
      if (FDR[s] > TH) {
        th <- j.np[s]
        break
      }
    }
    
    # Store the results for the current class in the `out` list
    out[[net]] <- out.jrf[out.jrf[, 2 + net] > th, seq(1, 2)]
  }
  
  return(out)
}

mynet <- myJRF_network(netout,out.perm,0.001)
mynet

```



```{r JRF inference}
#https://cran.r-project.org/src/contrib/Archive/JRF/
#install.packages("/home/francescoc/Downloads/JRF_0.1-4.tar.gz", repos = NULL, type = "source")
set.seed(1234)
tictoc::tic("JRF")
jrf_mat <- infer_networks(count_matrices, method="JRF")
execution_times[['JRF']] <- tictoc::toc(log = TRUE)$toc[[1]]

jrf_mat[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "JRF output")

```

#### Prepare the output

```{r}
jrf_list <- list()

importance_columns <- grep("importance", names(jrf_mat[[1]]), value = TRUE)

for (i in seq_along(importance_columns)) {
  # Select the 'gene1', 'gene2', and the current 'importance' column
  df <- jrf_mat[[1]][, c("gene1", "gene2", importance_columns[i])]
  
  # Rename the importance column to its original name (e.g., importance1, importance2, etc.)
  names(df)[3] <- importance_columns[i]
  
  # Add the data frame to the output list
  jrf_list[[i]] <- df
}

saveRDS(jrf_list, "./../analysis/jrf.RDS")

jrf_list[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "JRF output")
```

#### symmetrize Output and ROC

```{r JRF symmetrizeROC late}
jrf_wadj <- generate_adjacency(jrf_list, ground.truth = adjm)
sjrf_wadj <- symmetrize(jrf_wadj, weight_function = "mean")
plotROC(sjrf_wadj, adjm, plot_title = "ROC curve - JRF Late Integration")

sjrf_wadj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "JRF symmetrize output")
```

#### Generate Adjacency and Apply Cutoff

```{r JRF cutoff late}
sjrf_adj <- cutoff_adjacency(count_matrices = count_matrices,
                 weighted_adjm_list = sjrf_wadj, 
                 ground.truth = adjm,
                 n = 2,
                 method = "JRF")

sjrf_adj[[1]] %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "JRF adjacency")

```

#### Comparison with the Ground Truth

```{r JRF metrics late}
scores.jrf.all <- pscores(adjm, sjrf_adj)

scores.jrf.all$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
```


```{r JRF plotg late}
plots <- plotg(sjrf_adj)
```

```{r JRF consensus and score late}
consesusm <- create_consensus(sjrf_adj, method="vote")
consesusu <- create_consensus(sjrf_adj, method="union")

scores.jrf <- pscores(adjm, list(consesusm))
scoresu.jrf <- pscores(adjm, list(consesusu))

scores.jrf$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
scoresu.jrf$Statistics %>%
    datatable(extensions = 'Buttons',
              options = list(
                dom = 'Bfrtip',
                buttons = c('csv', 'excel'),
                scrollX = TRUE,
                pageLength = 10), 
              caption = "scores")
```

```{r JRF compare late, fig.width=8,fig.width=10}
source("compare_consensus.R")
ajm_compared <- compare_consensus(consesusm, adjm)
ajm_compared <- compare_consensus(consesusu, adjm)
```


# Method Comparison


## Time analysis

```{r plot_execution_times, echo=FALSE}
time_data <- data.frame(
  Method = names(execution_times),
  Time_in_Hours = unlist(execution_times) / 3600
)

time_data$Time_in_Minutes <- time_data$Time_in_Hours * 60

time_data <- time_data[order(time_data$Time_in_Hours), ]
time_data$Method <- factor(time_data$Method, levels = time_data$Method)

ggplot(time_data, aes(x = Method, y = Time_in_Hours)) +
  geom_bar(stat = 'identity') +
  geom_text(aes(label = sprintf("%.1f min", Time_in_Minutes)), 
            vjust = -0.5) +
  labs(title = 'Execution Time for Each Method', y = 'Time (Hours)', x = 'Method') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


## Statistics


```{r, fig.height=7, fig.width=5}
mplot <- list()

for (k in c("TPR", "FPR", "Precision", "F1", "Accuracy")) {
  mplot[[k]] <- data.frame(
    GENIE3_late = scores.genie3.late$Statistics[[k]],
    GENIE3_early = scores.genie3.early$Statistics[[k]],
    GRNBoost2_late = scores.grn.late$Statistics[[k]],
    GRNBoost2_early = scores.grn.early$Statistics[[k]],
    ZILGM_late = scores.zilgm.late$Statistics[[k]],
    ZILGM_early = scores.zilgm.early$Statistics[[k]],
    JRF = scores.jrf$Statistics[[k]]
  )
}

library(patchwork)  # For arranging plots into panels

# Convert mplot list into a long data frame for ggplot2
plot_data <- bind_rows(lapply(names(mplot), function(metric) {
  data.frame(
    Metric = metric,
    Method = names(mplot[[metric]]),
    Value = as.numeric(mplot[[metric]][1, ])
  )
}))

# Extract method groups for coloring and order the Method factor so JRF appears last
plot_data <- plot_data %>%
  mutate(Method_Group = case_when(
    grepl("GENIE3", Method) ~ "GENIE3",
    grepl("GRNBoost2", Method) ~ "GRNBoost2",
    grepl("ZILGM", Method) ~ "ZILGM",
    grepl("JRF", Method) ~ "JRF"
  )) %>%
  mutate(Method = factor(Method, levels = c(
    "GENIE3_early", "GENIE3_late", 
    "GRNBoost2_early", "GRNBoost2_late", 
    "ZILGM_early", "ZILGM_late", 
    "JRF"  # Ensure JRF is last
  )))

# Define color palette for each method group
method_colors <- c("GENIE3" = "darkblue", "GRNBoost2" = "darkgreen", "ZILGM" = "orange", "JRF" = "red")

# Create a separate ggplot for each metric, with y-axis limits set to 0-1 and no legend
plots <- lapply(unique(plot_data$Metric), function(metric) {
  # Determine if x-axis text should be displayed
  show_x_text <- metric %in% c("Accuracy", "F1")
  
  ggplot(plot_data %>% filter(Metric == metric), aes(x = Method, y = Value, fill = Method_Group)) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
    labs(title = metric, y = "Value", x = NULL) +
    scale_y_continuous(limits = c(0, 1)) +  # Set y-axis limits
    scale_fill_manual(values = method_colors) +
    theme_minimal() +
    theme(
      axis.text.x = if (show_x_text) element_text(angle = 45, hjust = 1) else element_blank(),
      axis.title.x = element_blank(),
      legend.position = "none"  # Remove legend from individual plots
    )
})

# Create a shared legend using one of the plots and add it at the bottom
# Assemble all plots in a 3x2 grid with the legend positioned at the bottom
final_plot <- (plots[[1]] | plots[[2]]) /
              (plots[[3]] | plots[[4]]) /
              (plots[[5]] + plot_spacer()) +
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

# Display the final plot
print(final_plot)
```



```{r}
sessionInfo()
```





